---
title: 'Part Two: Data Analysis'
author: "Henry Siegler"
date: '2022-04-07'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
```

```{r}
library(tidyverse)
library(here)
library(lubridate)
library(boot)
```

## The Data

```{r}
oscars <- read_csv(here("Oscars-demographics-DFE.csv"))
```

## Warm-Ups

__1. Which movie(s) won the most unique "Big 5" awards?__

```{r}
oscars %>% 
  distinct(movie, award) %>% 
  group_by(movie) %>% 
  count(sort = TRUE) %>% 
  filter(n == 3) %>% 
  pull(movie)
```

All the above movies won 3 unique "Big 5"awards.

__2. Of all the actresses who have won the Best Actress award, what is the most common first name?__

```{r}
name <- oscars %>% 
  distinct(person, .keep_all = TRUE) %>%
  filter(award == "Best Actress") %>%
  pull(person) %>%
  word(1) %>%
  table() %>% 
  sort(decreasing = TRUE)

head(name)
```
The most common first name is "Helen", with 3 occurences.

__3. What US State, or non-US country, has produced the most Oscar Winners (for the awards in the dataset)?__

```{r}
oscars$birthplace <- gsub("New York City", 
                          'New York City, Ny', 
                          oscars$birthplace)

state <- oscars %>% 
  distinct(person, .keep_all = TRUE) %>%
  pull(birthplace) %>%
  word(-1) %>%
  table() %>% 
  sort(decreasing = TRUE)

head(state)
```

Out of all US states or non-US countries, New York has produced the most Oscar winners, with 56 winners.

## Age and Gender

```{r}
oscars$date_of_birth[85] <- "5-Oct-1972"
oscars$date_of_birth[55] <- "18-Aug-1936"

modeldata <- oscars %>%
  separate(
    date_of_birth, 
    into = c("day", "month", "year"), 
    sep = "-") %>%
  mutate(
    day = as.numeric(day),
    year = as.numeric(year),
    year = case_when(year < 100 ~ 1900 + year,
                     TRUE ~ year),
    age = year_of_award - year,
    male = case_when(
      award == "Best Actress" |
      award == "Best Supporting Actress" ~ 0,
      award == "Best Actor" |
      award == "Best Supporting Actor" ~ 1,
      TRUE ~ 99)) %>%
  filter(male != 99)
```


```{r}
modeldata <- modeldata %>% 
  #create a year of award variable where the first year of the awards is 0
  mutate(year_award_0 = year_of_award - min(modeldata$year_of_award))


model <- lm(age ~ year_award_0 + male, data = modeldata)

summary(model)
```

A linear model was created to predict the age of the Oscars winner using the year they won and their gender, using a binary variable with 1 indicating male and 0 indicating female. The model predicts how the age of the average winner has changed over time as well as if male and female groups win awards at different ages.

From this model, we can say that there is moderate to strong evidence (p-value = 0.0517) that the average age of winners is changing over time and we have very strong evidence that males and female winners are of different average ages. The interaction term was removed from the model because it was not significant, telling us that the effect of the year of the award on the typical age of the award winner is similar for males and females.
The model estimates that the average age of winners increases by about three weeks (1/18th of a year), as the year of the award increases by 1 year. The model estimates that the average male winner is 9.5 years older than the average female winner.

## Bootstrapping

```{r}
notwhite <- oscars %>% 
  mutate(
    ethnicity = case_when(
      race_ethnicity != "White" ~ 1,
      TRUE ~ 0
    )) %>% 
  select(ethnicity)

set.seed(37)

boot <- bootstrap(notwhite, 1000)
boot <- mutate(boot, prop = map_dbl(strap, ~mean(data.frame(.x)$ethnicity)))

#Confidence Interval
lower = quantile(boot$prop, 0.025)
upper = quantile(boot$prop, 0.975)


p <- boot %>% 
  ggplot() +
  geom_density(aes(x = prop, y = ..density..), fill = "grey") 
  
mapdata <- ggplot_build(p)$data[[1]]

lower
upper
```

```{r}
p <-  p + 
  geom_area(aes(x=x, y=y), data = subset(mapdata, x < lower), fill = "red") + 
  geom_area(aes(x=x, y=y), data = subset(mapdata, x > upper), fill = "red") +
  ggtitle("Bootstraped Distribution of Proportions of Non-White Oscar Winners")

p
```

We are 95% confident that the true proportion of non-white oscars big 5 award winners is between 4.3% and 9.29%. The rejection regions for a hypothesis test for this proportion is shaded on the graph.